Dependent Object Types

### 2012-06-30 doc/semester/report.tex
wfe

Some subtyping transitivity is essential for soundness. Indeed, I show
how to construct a counterexample to preservation from any
counterexample to subtyping transitivity where the three involved
types $S, T, U$ are expressible within a realizable universe, though
the types themselves don't need to be realizable:

@@

The original DOT calculus has both object and function types. It
doesn't have built-in methods, but they can be expressed as functional
labels. On the other hand, Scala doesn't have a built-in function
type. Instead, it has both methods and fields, and encodes functions
as objects with an apply method. Here, I show that the DOT model as
originally designed is problematic.

@@

>>>>

Some subtyping transitivity is essential for soundness. Indeed, we show
how to construct a counterexample to preservation from any
counterexample to subtyping transitivity where the three involved
types $S, T, U$ are expressible within a realizable universe, though
the types themselves don't need to be realizable:

@@

The original DOT calculus has both object and function types. It
doesn't have built-in methods, but they can be expressed as functional
labels. On the other hand, Scala doesn't have a built-in function
type. Instead, it has both methods and fields, and encodes functions
as objects with an apply method. Here, we show that the DOT model as
originally designed is problematic.

@@

\section{Patches}

We have explored several patches to the calculus to deal with the
counterexamples presented above.

\subsection{Well-Formed and Expanding Types}\label{wfe}

We introduce a new judgement form for whether a type is well-formed
and expanding: $\Gamma \ts T \wfe$ if and only if $\Gamma \ts T \wf$
and $\exists \seq D$ such that $\Gamma \ts T \expand_z \seq D$. Then,
we replace all other uses of the $\wf$ judgment, including those
within the $\wf$ judgment inference rules, with uses of the $\wfe$
judgment.

We limit subtyping to $\wfe$ types. In fact, we make subtyping regular
with respect to $\wfe$: i.e., if $\Gamma \ts S <: T$, then $\Gamma \ts
S \wfe$ and $\Gamma \ts T \wfe$. Assuming a few lemmas, we are able to
formally prove that this modified subtyping judgment is
transitive. The most notable lemma we assume is the ``Galois
connection'' between subtyping and expansion: if $\Gamma \ts S \sub T
\spcomma S \expand_z \seq D_S \spcomma T \expand_z\seq D_T$, then
$\Gamma \envplus{z: S} \ts \seq{D_s \sub D_T}$.
